\pkg{casebase}: An Alternative Framework For Survival Analysis

### 2020-03-26 analysis/paper/paper.Rmd
AddedBrier
Still need to do:
- Formatting is a little off
- some plots are not optimized yet
-interpreting the brier score, casebase is pretty bad!

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = TRUE,
  comment = "#>",
  fig.path = "../figures/"
)
options(kableExtra.latex.load_packages = FALSE)
library(tidyverse)
library(magrittr)
library(survival)
library(casebase)
# devtools::load_all('../../../casebase')#testing
library(splines)
library(tibble)
library(glmnet)
library(data.table)
library(kableExtra)
library(pracma)
options(digits = 2, scipen = 999)
# set the seed for reproducible output
set.seed(1234)

@@

Before generating an absolute risk curve using regularized logistic regression, we will explore three options as a baseline comparison, beyond a simple Kaplan-meier curve. For all the following analyses in this case study, we will use 80% of the observations to train our models, while 20% will be used to generate absolute risk curves, all of which will be averaged to demonstrate each model's predictive ability. We will then compare each modeling using the IPA score (???NEED REFERENCE HERE???) We will examine three cox regression scenarios: sps only, aps only and all covariates excluding sps and aps. The original study compared these scores over a span of 180 days, which may not be comparable over the full 5.56 years.  Over 5.56 years, \ref{fig:baselines} demonstrates that ???BEST MODEL BASED ON IPA??? has the best calibration and discrimination with an IPA score of ???INSERT ACTUAL BEST SCORE??? . ??BEST MODEL?? will be used in the following comparison to regularized cox, regularized logistic regression and the unadjusted Kaplan-Meier curve.

@@

fullDataCox <- as.data.frame(cbind(as.numeric(y[, 2]), as.numeric(y[, 1]), x))
train_index <- sample(1:nrow(fullDataCox), 0.8 * nrow(fullDataCox))
test_index <- setdiff(1:nrow(fullDataCox), train_index)

train <- fullDataCox[train_index, ]
test <- fullDataCox[test_index, ]
coxSPS <- survival::coxph(Surv(time = V1, event = V2) ~ sps, data = train)
abcoxSPS <- survival::survfit(coxSPS, newdata = test)

coxAPS <- survival::coxph(Surv(time = V1, event = V2) ~ aps, data = train)
abcoxAPS <- survival::survfit(coxAPS, newdata = test)

coxFull <- survival::coxph(Surv(time = V1, event = V2) ~ . - sps - aps, data = train)
abcoxFull <- survival::survfit(coxAPS, newdata = test)
# library(prodlim)

KM <- survival::coxph(Surv(time = V1, event = V2) ~ 1, data = test)
abKM <- survival::survfit(KM)
# abKM <- prodlim(Hist(V1,V2)~1,data=test) minimal difference
abKMcompare <- cbind(abKM$time[abKM$time <= 1631], 1 - abKM$surv[abKM$time <= 1631])
colnames(abKMcompare) <- c("Time", "KM")

baseCoxComparisonData <- cbind(abcoxSPS$time[abcoxSPS$time <= 1631], 1 - rowMeans(abcoxSPS$surv[abcoxSPS$time <= 1631, ]), 1 - rowMeans(abcoxAPS$surv[abcoxSPS$time <= 1631, ]), 1 - abKM$surv, 1 - rowMeans(abcoxFull$surv[abcoxSPS$time <= 1631, ]))
colnames(baseCoxComparisonData) <- c("Time", "SPS", "APS", "Unadjusted", "Full")
```

@@

@@

For our penalized logistic regression model, we opted for a logistic function of time, resulting in a Weibull distribution assumption. As casebase makes use of the glmnet package, we interact with the fitsmoothhazard.fit function using a matrix interface, where y contains the time and event variables and x contains all other variables we would like to include in the model. For the SUPPORT dataset, all factor variables were converted into dummy variables. Here, we used lasso penalization by setting alpha to 1, leading to our smooth hazard function, depicted in the ??? curve in \ref{fig:cs3Final}.
??? I AM STILL PENALIZING TIME ???

We then take our hazard and use it to predict the absolute risk of each individual in our test set, over time. These individuals are then average resulting in the ???curve in \ref{fig:cs3Final}. ???SAME BUG IS BACK???

@@

# fitting a cox model using regular estimation, however we will not keep it.
# this is used more as an object place holder.
coxNet <- survival::coxph(Surv(time = d.time, event = death) ~ ., data = cleanCoxData)

@@

estimatescoxNet$Model <- "CoxNet"
colnames(estimatescoxNet) <- c("term", "estimate", "model")

@@

@@

cbFull <- casebase::fitSmoothHazard.fit(x, y, family = "glmnet", time = "d.time", event = "death", formula_time = ~ log(d.time), alpha = 1, ratio = 10, standardize = TRUE)
```

```{r CBabsolute, echo = TRUE, eval = eval_cs3}
# Estimating the absolute risk curve using the newData parameter.
abCbFull <- casebase::absoluteRisk(cbFull, time = seq(1, max(abKMcompare[, 1]), 25), newdata = newData, s = "lambda.1se", method = c("numerical"))
```

@@

```{r include=TRUE,eval=eval_cs3,fig.cap="\\label{fig:baselines} HERES SOME TEXT ABT IT"}
ggplot(as.data.frame(ab), aes(Time)) +
  geom_line(data = as.data.frame(baseCoxComparisonData), mapping = aes(y = Full, colour = "Cox"), lwd = 2) +
  geom_line(aes(y = `Absolute risk`, colour = Model), lwd = 2) +
  geom_step(data = as.data.frame(abKMcompare), mapping = aes(x = Time, y = KM, color = "Unadjusted"), lwd = 2) +
  labs(y = "Probability of Death", x = "Survival time (Days)")
```

@@

```{r coefplots, eval=eval_cs3}

@@

estimatescb <- setDT(as.data.frame(coef(cbFull)[-c(1, 2), 1]), keep.rownames = TRUE)[]
# estimatescb$std.error<-0
estimatescb$Model <- "casebase"
colnames(estimatescb) <- c("term", "estimate", "model")

estimatescoxFull <- setDT(as.data.frame(coef(coxFull)), keep.rownames = TRUE)[]
# estimatescoxFull$std.error<-0
estimatescoxFull$Model <- "CoxFull"
colnames(estimatescoxFull) <- c("term", "estimate", "model")

lolliplotDots <- rbind(estimatescb, estimatescoxNet, estimatescoxFull)
lolliplotDots$conf.low <- lolliplotDots$estimate
lolliplotDots$conf.high <- lolliplotDots$estimate
lolliplotDots$conf.high[lolliplotDots$estimate < 0] <- 0
lolliplotDots$conf.low[lolliplotDots$estimate > 0] <- 0

dwplot(lolliplotDots)
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = TRUE,
  comment = "#>",
  fig.path = "../figures/"
)
options(kableExtra.latex.load_packages = FALSE)
library(tidyverse)
library(magrittr)
library(survival)
library(casebase)
# devtools::load_all('../../../casebase')#testing
library(splines)
library(tibble)
library(glmnet)
library(data.table)
library(kableExtra)
library(pracma)
library(riskRegression)
library(prodlim)
options(digits = 2, scipen = 999)
# set the seed for reproducible output
set.seed(1234)

@@

Before generating an absolute risk curve using regularized logistic regression, we will explore three options as a baseline comparison, beyond a simple Kaplan-meier curve. For all the following analyses in this case study, we will use 80% of the observations to train our models, while 20% will be used to generate absolute risk curves, all of which will be averaged to demonstrate each model's predictive ability. We will then compare each modeling using the brier score [@graf1999ass] We will examine three cox regression scenarios: sps only, aps only and all covariates excluding sps and aps.

The original study compared these scores over a span of 180 days, which may not be comparable over the full 5.56 years.  Over 5.56 years, \ref{fig:baselines} demonstrates that all the cox models have similar, if not identical average predictions. Figure \ref{fig:brier1} examines the brier scores as they change over time. As the full model has the lowest brier score over all, it will be used in the following comparison to regularized cox, regularized logistic regression and the unadjusted Kaplan-Meier curve.

@@

fullDataCox <- as.data.frame(cbind(as.numeric(y[, 2]), as.numeric(y[, 1]), x))
names(fullDataCox)[names(fullDataCox) == 'V1'] <- 'time'
names(fullDataCox)[names(fullDataCox) == 'V2'] <- 'event'
train_index <- sample(1:nrow(fullDataCox), 0.8 * nrow(fullDataCox))
test_index <- setdiff(1:nrow(fullDataCox), train_index)

train <- fullDataCox[train_index, ]
test <- fullDataCox[test_index, ]
coxSPS <- survival::coxph(Surv(time =time , event = event) ~ sps, data = train, x = TRUE)
abcoxSPS <- survival::survfit(coxSPS, newdata = test)

coxAPS <- survival::coxph(Surv(time =time , event = event) ~ aps, data = train, x = TRUE)
abcoxAPS <- survival::survfit(coxAPS, newdata = test)
coxfullForumla<-as.formula(paste("Surv(time = time, event = event) ~ age + sexmale + slos + dzgroupCHF + 
    dzgroupCirrhosis + `dzgroupColon Cancer` + dzgroupComa + 
    dzgroupCOPD + `dzgroupLung Cancer` + `dzgroupMOSF w/Malig` + 
    dzclassCancer + dzclassComa + `dzclassCOPD/CHF/Cirrhosis` + 
    num.co + edu + scoma + avtisst + raceasian + raceblack + 
    racehispanic + raceother + racewhite + hday + diabetes + 
    dementia + cano + cayes + meanbp + wblc + hrt + resp + temp + 
    pafi + alb + bili + crea + sod + ph + glucose + bun + urine + 
    adlp + adlsc "))
coxFull <- survival::coxph(coxfullForumla, data = train, x = TRUE)
abcoxFull <- survival::survfit(coxFull, newdata = test)
# library(prodlim)

KM <- survival::coxph(Surv(time = time, event = event) ~ 1, data = test)
abKM <- survival::survfit(KM)
# abKM <- prodlim(Hist(,event)~1,data=test) minimal difference
abKMcompare <- cbind(abKM$time[abKM$time <= 1631], 1 - abKM$surv[abKM$time <= 1631])
colnames(abKMcompare) <- c("Time", "KM")

baseCoxComparisonData <- cbind(abcoxSPS$time[abcoxSPS$time <= 1631], 1 - rowMeans(abcoxSPS$surv[abcoxSPS$time <= 1631, ]), 1 - rowMeans(abcoxAPS$surv[abcoxAPS$time <= 1631, ]), 1 - rowMeans(abcoxFull$surv[abcoxFull$time <= 1631, ]))

colnames(baseCoxComparisonData) <- c("Time", "SPS", "APS","Full")
```

@@

```{r brierpt1, echo=FALSE,eval=eval_cs3,fig.cap="\\label{fig:brier1} Brier scores over time. Full model performs best.",echo=FALSE}

times <- sort(unique(test$time))
brierAPS <- Score(list("PredictionModel" = coxAPS), data = test, formula = Hist(time, event != 0) ~ 1, summary = "ipa", se.fit = 0L, metrics = "brier", contrasts = FALSE, times = times)
brierSPS <- Score(list("PredictionModel" = coxSPS), data = test, formula = Hist(time, event != 0) ~ 1, summary = "ipa", se.fit = 0L, metrics = "brier", contrasts = FALSE, times = times)

coxFull$coefficients[which(is.na(coxFull$coefficients))]=0
brierFull <- Score(list("PredictionModel" = coxFull), data = test, formula = Surv(time, event != 0) ~ 1, summary = "ipa", se.fit = 0L, metrics = "brier", contrasts = FALSE, times = times)

brierPlotter<-rbind(
      data.frame(brierScore = brierFull$Brier$score$Brier[brierFull$Brier$score$model == "PredictionModel"],times=times,model="cox Full"),
      data.frame(brierScore = brierAPS$Brier$score$Brier[brierAPS$Brier$score$model == "PredictionModel"],times=times,model="cox APS only"),
            data.frame(brierScore = brierSPS$Brier$score$Brier[brierSPS$Brier$score$model == "PredictionModel"],times=times,model="cox SPS only")
      )

ggplot(data = brierPlotter, mapping = aes(x = times, y = brierScore, col = model)) +
  geom_line() +
  xlab("Times") +
  ylab("Brier-Score")+
  ggtitle("Brier scores over time")
```

@@

For our penalized logistic regression model, we opted for a logistic function of time, resulting in a Weibull distribution assumption. As casebase makes use of the glmnet package, we interact with the fitsmoothhazard.fit function using a matrix interface, where y contains the time and event variables and x contains all other variables we would like to include in the model. For the SUPPORT dataset, all factor variables were converted into dummy variables. Here, we used lasso penalization by setting alpha to 1, leading to our smooth hazard function.
 We then take our hazard and use it to predict the absolute risk of each individual in our test set, over time. These individuals are then average resulting in the smooth absolute risk curve in Figure \ref{fig:cs3Final}.

**note that I am still penalizing time, there's a bug I did not want to deal with when I try to use `penalty.factor` in fitsmoothhazard . If this is still here, i have not fixed it.**

@@

# fitting a cox model using regular estimation, however we will not keep it.
# this is used more as an object place holder.
coxNet <- survival::coxph(Surv(time = d.time, event = death) ~ ., data = cleanCoxData,x=TRUE)

@@

estimatescoxNet$Model <- "Penalized Cox"
colnames(estimatescoxNet) <- c("term", "estimate", "model")

@@

brierCoxNet <- Score(list("PredictionModel" = coxNet), data = as.data.frame(cbind(test[,c(1,2)],newData[, nonzero_covariate_cox$X1])), formula = Hist(time, event != 0) ~ 1, summary = "ipa", se.fit = 0L, metrics = "brier", contrasts = FALSE, times = times)

@@

cbFull <- casebase::fitSmoothHazard.fit(x, y, family = "glmnet", time = "d.time", event = "death", formula_time = ~ log(d.time), alpha = 1, ratio = 100, standardize = TRUE)
```

```{r CBabsolute, echo = TRUE, eval = eval_cs3}
# Estimating the absolute risk curve using the newData parameter.
abCbFull <- casebase::absoluteRisk(cbFull, time = times, newdata = newData, s = "lambda.1se", method = c("numerical"))
```

@@

```{r include=TRUE,echo=FALSE,eval=eval_cs3,fig.cap="\\label{fig:cs3Final} HERES SOME TEXT ABT IT"}
ggplot(as.data.frame(ab), aes(Time)) +
  geom_line(data = as.data.frame(baseCoxComparisonData), mapping = aes(y = Full, colour = "Cox"), lwd = 2) +
  geom_line(aes(y = `Absolute risk`, colour = Model), lwd = 2) +
  geom_step(data = as.data.frame(abKMcompare), mapping = aes(x = Time, y = KM, color = "Unadjusted"), lwd = 2) +
  labs(y = "Probability of Death", x = "Survival time (Days)")
```

@@

```{r finalBrierplot, echo=FALSE, eval=eval_cs3}
brierScoreKMCens <- function(predSurvs, times, newData) {
  # inverse probability censoring weights
  # Note that we take probabilities right before the drops.

  #DIFFERENCES START HERE
  fitCens <- prodlim::prodlim(Hist(time, event != 0) ~ 1, newData, reverse = TRUE)
  IPCW.subject.times <- prodlim::predictSurvIndividual(fitCens, lag = 1) # G(t-|X)
  #DIFFERENCES END HERE

  # Empty matrix that will be filled in with the following loop
  Score <- matrix(NA, nrow(predSurvs), ncol(predSurvs))
  IPCWMatrix <- matrix(NA, nrow(predSurvs), ncol(predSurvs))
  # for each point in time we have predicted
  for (i in 1:length(times)) {
    # get number of censored individuals so long as their survival time is less than times[i]
    # these individuals do not have an effect on right-censored brier score.
    CensBefore <- newData$event == 0 & newData$time < times[i]
    # y encompasses all survival times larger than t[i] with a 1, 0 otherwise
    y <- drop(t(newData$time > times[i]))
    # above permits the two parts of right-censored brier score to be calculated, without IPCW, in one line
    Score[i, ] <- (y - predSurvs[i, ])^2

    # Generate IPCWMatrix
    IPCWMatrix[i, y == 0] <- IPCW.subject.times[y == 0] # G(t-|X) filled in corresponding positions
    #DIFFERENCES START HERE
    IPCW.time <- predict(fitCens, newdata = newData, times = times[i], level.chaos = 1, mode = "matrix", type = "surv", lag = 1)
    #DIFFERENCES END HERE
    IPCWMatrix[i, y == 1] <- IPCW.time # G(t) filled, same value, for remaining positions.
    # above calculated individuals who don't have an effect on brier score at specfied time. they're scores are set to 0.
    Score[i, CensBefore] <- 0
  }

  # apply IPCW to all scores
  Err <- Score / IPCWMatrix

  # Average curve demonstrating right-censored brier averaged over test-set, for each time of interest.
  Err <- apply(Err, 1, mean)
  return(Err)
}
absRiskcb <-abCbFull
absRiskcb[, -c(1)] <- 1 - absRiskcb[, -c(1)] # make it survival probabilities
absRiskcb <- absRiskcb[-c(1), -c(1)] # remove extra time at 0. Also remove the first column (times) as it matches perfectly now

kmCensoredBrierScores <- brierScoreKMCens(predSurvs = absRiskcb, times = times, newData = data.frame(time=test$time,event=test$event))

finalBrierDataToPlot<-rbind(
  data.frame(brierScore = brierCoxNet$Brier$score$Brier[brierCoxNet$Brier$score$model == "PredictionModel"],times=times,model="Penalized Cox"),
   data.frame(brierScore = brierFull$Brier$score$Brier[brierFull$Brier$score$model == "PredictionModel"],times=times,model="cox"),
      rbind(data.frame(brierScore=kmCensoredBrierScores,times=times,model="Penalized logistic"))

  )

ggplot(data = finalBrierDataToPlot, mapping = aes(x = times, y = brierScore, col = model)) +
  geom_line() +
  xlab("Times") +
  ylab("Brier-Score")+
  ggtitle("Brier scores over time")

```

```{r coefplots,echo=FALSE, eval=eval_cs3}

@@

estimatescb <- setDT(as.data.frame(coef(cbFull)[-c(1, 2), 1]), keep.rownames = TRUE)[]
# estimatescb$std.error<-0
estimatescb$Model <- "Penalized logistic"
colnames(estimatescb) <- c("term", "estimate", "model")

estimatescoxFull <- setDT(as.data.frame(coef(coxFull)), keep.rownames = TRUE)[]
# estimatescoxFull$std.error<-0
estimatescoxFull$Model <- "Cox"
colnames(estimatescoxFull) <- c("term", "estimate", "model")

lolliplotDots <- rbind(estimatescb, estimatescoxNet, estimatescoxFull)
lolliplotDots$conf.low <- lolliplotDots$estimate
lolliplotDots$conf.high <- lolliplotDots$estimate
lolliplotDots$conf.high[lolliplotDots$estimate < 0] <- 0
lolliplotDots$conf.low[lolliplotDots$estimate > 0] <- 0
lolliplotDots$term=gsub( '\`', "", as.character(lolliplotDots$term))
dwplot(lolliplotDots)+
  theme(axis.text.y = element_text(size = 7.5, angle = 0, hjust = 1, vjust = 0),
        strip.text.x = element_blank(),
      strip.background = element_rect(colour="white", fill="white"),
      legend.position=c(.855,0.15),legend.text=element_text(size=7.5)
      )
```

### 2020-03-26 analysis/paper/paper.tex
Auto stash before merge of "write-popTime" and "origin/master"
AddedBrier
Still need to do:
- Formatting is a little off
- some plots are not optimized yet
-interpreting the brier score, casebase is pretty bad!

Before generating an absolute risk curve using regularized logistic
regression, we will explore three options as a baseline comparison,
beyond a simple Kaplan-meier curve. For all the following analyses in
this case study, we will use 80\% of the observations to train our
models, while 20\% will be used to generate absolute risk curves, all of
which will be averaged to demonstrate each model's predictive ability.
We will then compare each modeling using the IPA score (???NEED
REFERENCE HERE???) We will examine three cox regression scenarios: sps
only, aps only and all covariates excluding sps and aps. The original
study compared these scores over a span of 180 days, which may not be
comparable over the full 5.56 years. Over 5.56 years,
\ref{fig:baselines} demonstrates that ???BEST MODEL BASED ON IPA??? has
the best calibration and discrimination with an IPA score of ???INSERT
ACTUAL BEST SCORE??? . ??BEST MODEL?? will be used in the following
comparison to regularized cox, regularized logistic regression and the
unadjusted Kaplan-Meier curve.

@@

@@

For our penalized logistic regression model, we opted for a logistic
function of time, resulting in a Weibull distribution assumption. As
casebase makes use of the glmnet package, we interact with the
fitsmoothhazard.fit function using a matrix interface, where y contains
the time and event variables and x contains all other variables we would
like to include in the model. For the SUPPORT dataset, all factor
variables were converted into dummy variables. Here, we used lasso
penalization by setting alpha to 1, leading to our smooth hazard
function, depicted in the ??? curve in \ref{fig:cs3Final}. ??? I AM
STILL PENALIZING TIME ???

We then take our hazard and use it to predict the absolute risk of each
individual in our test set, over time. These individuals are then
average resulting in the ???curve in \ref{fig:cs3Final}. ???SAME BUG IS
BACK???

@@

\begin{CodeInput}
R> 
R> cbFull <- casebase::fitSmoothHazard.fit(x, y, family = "glmnet", time = "d.time", event = "death", formula_time = ~ log(d.time), alpha = 1, ratio = ratio, standardize = TRUE)
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> # Estimating the absolute risk curve using the newData parameter.
R> abCbFull <- casebase::absoluteRisk(cbFull, time = seq(1, max(abKMcompare[, 1]), 25), newdata = newData, s = "lambda.1se", method = c("numerical"))
\end{CodeInput}
\end{CodeChunk}

\begin{CodeInput}
R> ggplot(as.data.frame(ab), aes(Time)) +
R+   geom_line(data = as.data.frame(baseCoxComparisonData), mapping = aes(y = Full, colour = "Cox"), lwd = 2) +
R+   geom_line(aes(y = `Absolute risk`, colour = Model), lwd = 2) +
R+   geom_step(data = as.data.frame(abKMcompare), mapping = aes(x = Time, y = KM, color = "Unadjusted"), lwd = 2) +
R+   labs(y = "Probability of Death", x = "Survival time (Days)")
\end{CodeInput}
\begin{figure}

@@

\caption{\label{fig:baselines} HERES SOME TEXT ABT IT}\label{fig:unnamed-chunk-2}
\end{figure}
\end{CodeChunk}

@@

\begin{CodeInput}
R> 
R> library(dotwhisker)
R> library(broom)
R> library(dplyr)
R> 
R> estimatescb <- setDT(as.data.frame(coef(cbFull)[-c(1, 2), 1]), keep.rownames = TRUE)[]
R> # estimatescb$std.error<-0
R> estimatescb$Model <- "casebase"
R> colnames(estimatescb) <- c("term", "estimate", "model")
R> 
R> estimatescoxFull <- setDT(as.data.frame(coef(coxFull)), keep.rownames = TRUE)[]
R> # estimatescoxFull$std.error<-0
R> estimatescoxFull$Model <- "CoxFull"
R> colnames(estimatescoxFull) <- c("term", "estimate", "model")
R> 
R> lolliplotDots <- rbind(estimatescb, estimatescoxNet, estimatescoxFull)
R> lolliplotDots$conf.low <- lolliplotDots$estimate
R> lolliplotDots$conf.high <- lolliplotDots$estimate
R> lolliplotDots$conf.high[lolliplotDots$estimate < 0] <- 0
R> lolliplotDots$conf.low[lolliplotDots$estimate > 0] <- 0
R> 
R> dwplot(lolliplotDots)
\end{CodeInput}

@@

This report was generated on 2020-03-06 08:01:48 using the following
computational environment and dependencies:

@@

\begin{CodeOutput}
#> R version 3.6.1 (2019-07-05)
#> Platform: x86_64-w64-mingw32/x64 (64-bit)
#> Running under: Windows 10 x64 (build 18363)
#> 
#> Matrix products: default
#> 
#> locale:
#> [1] LC_COLLATE=English_United States.1252 
#> [2] LC_CTYPE=English_United States.1252   
#> [3] LC_MONETARY=English_United States.1252
#> [4] LC_NUMERIC=C                          
#> [5] LC_TIME=English_United States.1252    
#> 
#> attached base packages:
#> [1] splines   stats     graphics  grDevices utils     datasets  methods  
#> [8] base     
#> 
#> other attached packages:
#>  [1] broom_0.5.5         dotwhisker_0.5.0    pracma_2.2.9       
#>  [4] kableExtra_1.1.0    data.table_1.12.8   glmnet_3.0-2       
#>  [7] Matrix_1.2-17       casebase_0.2.1.9001 survival_3.1-8     
#> [10] magrittr_1.5        forcats_0.5.0       stringr_1.4.0      
#> [13] dplyr_0.8.3         purrr_0.3.3         readr_1.3.1        
#> [16] tidyr_1.0.2         tibble_2.1.3        ggplot2_3.3.0      
#> [19] tidyverse_1.3.0    
#> 
#> loaded via a namespace (and not attached):
#>  [1] httr_1.4.1        VGAM_1.1-2        jsonlite_1.6.1    viridisLite_0.3.0
#>  [5] foreach_1.4.8     modelr_0.1.6      assertthat_0.2.1  stats4_3.6.1     
#>  [9] ggstance_0.3.3    cellranger_1.1.0  yaml_2.2.1        pillar_1.4.3     
#> [13] backports_1.1.5   lattice_0.20-38   glue_1.3.1        digest_0.6.25    
#> [17] rvest_0.3.5       colorspace_1.4-1  plyr_1.8.6        htmltools_0.4.0  
#> [21] pkgconfig_2.0.3   haven_2.2.0       scales_1.1.0      webshot_0.5.2    
#> [25] mgcv_1.8-28       rticles_0.14      generics_0.0.2    farver_2.0.3     
#> [29] withr_2.1.2       cli_2.0.2         crayon_1.3.4      readxl_1.3.1     
#> [33] evaluate_0.14     fs_1.3.2          fansi_0.4.1       nlme_3.1-140     
#> [37] xml2_1.2.2        tools_3.6.1       hms_0.5.3         lifecycle_0.1.0  
#> [41] munsell_0.5.0     reprex_0.3.0      compiler_3.6.1    rlang_0.4.5      
#> [45] grid_3.6.1        iterators_1.0.12  rstudioapi_0.11   labeling_0.3     
#> [49] rmarkdown_2.1     gtable_0.3.0      codetools_0.2-16  DBI_1.1.0        
#> [53] R6_2.4.1          lubridate_1.7.4   knitr_1.28        shape_1.4.4      
#> [57] stringi_1.4.6     Rcpp_1.0.3        vctrs_0.2.3       dbplyr_1.4.2     
#> [61] tidyselect_1.0.0  xfun_0.12
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeOutput}
#> Local:    master C:/Users/jesse/Documents/gits/cbpaper
#> Remote:   master @ origin (https://github.com/Jesse-Islam/cbpaper.git)
#> Head:     [6682dfb] 2020-03-06: NOT COMPLETE
\end{CodeOutput}
\end{CodeChunk}

>>>>

Before generating an absolute risk curve using regularized logistic
regression, we will explore three options as a baseline comparison,
beyond a simple Kaplan-meier curve. For all the following analyses in
this case study, we will use 80\% of the observations to train our
models, while 20\% will be used to generate absolute risk curves, all of
which will be averaged to demonstrate each model's predictive ability.
We will then compare each modeling using the brier score
\citep{graf1999ass} We will examine three cox regression scenarios: sps
only, aps only and all covariates excluding sps and aps.

The original study compared these scores over a span of 180 days, which
may not be comparable over the full 5.56 years. Over 5.56 years,
\ref{fig:baselines} demonstrates that all the cox models have similar,
if not identical average predictions. Figure \ref{fig:brier1} examines
the brier scores as they change over time. As the full model has the
lowest brier score over all, it will be used in the following comparison
to regularized cox, regularized logistic regression and the unadjusted
Kaplan-Meier curve.

@@

{\centering \includegraphics{../figures/brierpt1-1}

\caption{\label{fig:brier1} Brier scores over time. Full model performs best.}\label{fig:brierpt1}
\end{figure}
\end{CodeChunk}

@@

For our penalized logistic regression model, we opted for a logistic
function of time, resulting in a Weibull distribution assumption. As
casebase makes use of the glmnet package, we interact with the
fitsmoothhazard.fit function using a matrix interface, where y contains
the time and event variables and x contains all other variables we would
like to include in the model. For the SUPPORT dataset, all factor
variables were converted into dummy variables. Here, we used lasso
penalization by setting alpha to 1, leading to our smooth hazard
function. We then take our hazard and use it to predict the absolute
risk of each individual in our test set, over time. These individuals
are then average resulting in the smooth absolute risk curve in Figure
\ref{fig:cs3Final}.

\textbf{note that I am still penalizing time, there's a bug I did not
want to deal with when I try to use \texttt{penalty.factor} in
fitsmoothhazard . If this is still here, i have not fixed it.}

@@

\begin{CodeInput}
R> 
R> cbFull <- casebase::fitSmoothHazard.fit(x, y, family = "glmnet", time = "d.time", event = "death", formula_time = ~ log(d.time), alpha = 1, ratio = 100, standardize = TRUE)
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeInput}
R> # Estimating the absolute risk curve using the newData parameter.
R> abCbFull <- casebase::absoluteRisk(cbFull, time = times, newdata = newData, s = "lambda.1se", method = c("numerical"))
\end{CodeInput}
\end{CodeChunk}

@@

\caption{\label{fig:cs3Final} HERES SOME TEXT ABT IT}\label{fig:unnamed-chunk-2}
\end{figure}
\end{CodeChunk}

@@

\begin{center}\includegraphics{../figures/finalBrierplot-1} \end{center}

@@

This report was generated on 2020-03-26 16:26:24 using the following
computational environment and dependencies:

@@

\begin{CodeOutput}
#> R version 3.6.1 (2019-07-05)
#> Platform: x86_64-w64-mingw32/x64 (64-bit)
#> Running under: Windows 10 x64 (build 18363)
#> 
#> Matrix products: default
#> 
#> locale:
#> [1] LC_COLLATE=English_United States.1252 
#> [2] LC_CTYPE=English_United States.1252   
#> [3] LC_MONETARY=English_United States.1252
#> [4] LC_NUMERIC=C                          
#> [5] LC_TIME=English_United States.1252    
#> 
#> attached base packages:
#> [1] splines   stats     graphics  grDevices utils     datasets  methods  
#> [8] base     
#> 
#> other attached packages:
#>  [1] broom_0.5.5               dotwhisker_0.5.0         
#>  [3] prodlim_2019.11.13        riskRegression_2020.02.05
#>  [5] pracma_2.2.9              kableExtra_1.1.0         
#>  [7] data.table_1.12.8         glmnet_3.0-2             
#>  [9] Matrix_1.2-17             casebase_0.2.1.9001      
#> [11] survival_3.1-8            magrittr_1.5             
#> [13] forcats_0.5.0             stringr_1.4.0            
#> [15] dplyr_0.8.3               purrr_0.3.3              
#> [17] readr_1.3.1               tidyr_1.0.2              
#> [19] tibble_2.1.3              ggplot2_3.3.0            
#> [21] tidyverse_1.3.0          
#> 
#> loaded via a namespace (and not attached):
#>  [1] nlme_3.1-140        fs_1.3.2            cmprsk_2.2-9       
#>  [4] lubridate_1.7.4     rms_5.1-4           webshot_0.5.2      
#>  [7] RColorBrewer_1.1-2  httr_1.4.1          numDeriv_2016.8-1.1
#> [10] tools_3.6.1         backports_1.1.5     R6_2.4.1           
#> [13] rpart_4.1-15        Hmisc_4.3-1         DBI_1.1.0          
#> [16] mgcv_1.8-28         colorspace_1.4-1    nnet_7.3-12        
#> [19] withr_2.1.2         gridExtra_2.3       tidyselect_1.0.0   
#> [22] compiler_3.6.1      quantreg_5.54       cli_2.0.2          
#> [25] rvest_0.3.5         htmlTable_1.13.3    SparseM_1.78       
#> [28] xml2_1.2.5          sandwich_2.5-1      labeling_0.3       
#> [31] checkmate_2.0.0     scales_1.1.0        mvtnorm_1.1-0      
#> [34] polspline_1.1.17    digest_0.6.25       foreign_0.8-71     
#> [37] rmarkdown_2.1       base64enc_0.1-3     rticles_0.14       
#> [40] jpeg_0.1-8.1        pkgconfig_2.0.3     htmltools_0.4.0    
#> [43] dbplyr_1.4.2        htmlwidgets_1.5.1   rlang_0.4.5        
#> [46] readxl_1.3.1        rstudioapi_0.11     VGAM_1.1-2         
#> [49] farver_2.0.3        shape_1.4.4         generics_0.0.2     
#> [52] zoo_1.8-7           jsonlite_1.6.1      acepack_1.4.1      
#> [55] Formula_1.2-3       Rcpp_1.0.3          munsell_0.5.0      
#> [58] fansi_0.4.1         lifecycle_0.2.0     multcomp_1.4-12    
#> [61] stringi_1.4.6       yaml_2.2.1          MASS_7.3-51.4      
#> [64] plyr_1.8.6          ggstance_0.3.3      grid_3.6.1         
#> [67] crayon_1.3.4        lattice_0.20-38     haven_2.2.0        
#> [70] hms_0.5.3           timereg_1.9.4       knitr_1.28         
#> [73] pillar_1.4.3        codetools_0.2-16    stats4_3.6.1       
#> [76] reprex_0.3.0        glue_1.3.1          evaluate_0.14      
#> [79] latticeExtra_0.6-29 modelr_0.1.6        vctrs_0.2.4        
#> [82] png_0.1-7           foreach_1.4.8       MatrixModels_0.4-1 
#> [85] cellranger_1.1.0    gtable_0.3.0        assertthat_0.2.1   
#> [88] xfun_0.12           viridisLite_0.3.0   iterators_1.0.12   
#> [91] cluster_2.1.0       lava_1.6.7          TH.data_1.0-10
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeOutput}
#> Local:    master C:/Users/jesse/Documents/gits/cbpaper
#> Remote:   master @ origin (https://github.com/Jesse-Islam/cbpaper.git)
#> Head:     [a01a2a8] 2020-03-06: Fixed lollipop plot
\end{CodeOutput}
\end{CodeChunk}
